---
title: "Lab 3 DRAFT"
author: "Siddhartha Jakkamreddy, Neha Kumar, Brian Musisi"
date: "11/14/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=7, fig.height=3.5, fig.align='center')
library(purrr)
library(car)
library(ggplot2)
library(maps)
library(Hmisc)
library(stringr)
library(reshape2)
library(dplyr)
library(plyr)
library(tidyr)
library(lubridate)
library(zoo)
library(corrplot)
library(stargazer)
```

## Introduction

The motivation for this analysis is to determine the factors that lead to crime rate in North Carolina counties in 1980. We are assuming the role of data scientists for a political campaign around the same era within North Carolina to determine methods that can be employed to reduce the crime rate. Note, that this requires our analysis to look for causal variables so we can provide concrete and actionable resolutions.

## Initial EDA

```{r}
crime = read.csv("crime_v2.csv", header = TRUE)
str(crime)
```
Here, we see something interesting. prbconv is a factor due to a rogue ` character being added to the bottom of the file. When we view the dataframe, we actually see that this rogue tick mark has also introduced 6 null values in the file. We take steps to remove these records from the file to clean our dataset. 

```{r}
crime <- crime[!is.na(as.numeric(as.character(crime$prbconv))),]
crime$prbconv <- as.numeric(as.character(crime$prbconv))
str(crime)
summary(crime)
```

Something we notice here is that both the probability of arrest and probability of conviction have at least 1 record that is over 1. Thinking through this further, this is not impossible. Multiple people can participate in a crime together, leading to multiple arrests and/or convictions per criminal offense and so this anomaly may be a product of the operationalization of the particular variables. Thus, we will not discard this variable.

Our data file is now clean and ready for further analysis. 

## Model Building Process

```{r, fig.width=10,fig.height=11}
crime %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram()
```

From the above, the distribution of most of the numeric variables presented are approximately normal, with the exception of prbconv, pctymyle and polpc that stand out as non-normal.  We now take a look at a scatterplot matrix to examine the bivariate relationships embedded within this dataset.

Looking at the variables, we have a few initial thoughts. There is suspected collinearity between wage variables and tax revenue per capita. We also note that higher tax per capita allows counties to spend more money on police forces, possibly having an effect on the police per capita. Lastly, density could have a confounding effect on per capita variables. Assume we have 2 counties, ceteris paribis, differing only in population density. This means that the per capita measurements for the more densely populated county will be lower than the more sparsely populated county.

We create a correlation matrix to get a high level overview of the correlations between variables.

```{r}
corr_crime = cor(crime[, c(-1,-2,-11,-12,-13)])  ### removing the urban, west and central variables due to low variability for a correlation plot
corrplot(corr_crime, method = "ellipse")
```

### Examining density

The first relationships we decide to investigate further are the density variables against the binary urban, west and central variables. From the correlation matrix and our intuition, we expect that urban areas are more dense. 

```{r}
crime %>%
  ggplot(aes(density)) +
    facet_wrap(~ urban) +
    geom_histogram() +
    ggtitle("Non Urban (0) vs Urban Density (1)")

crime %>%
  ggplot(aes(density)) +
    facet_wrap(~ west) +
    geom_histogram() + 
    ggtitle("Non West (0) vs West Density (1)")

crime %>%
  ggplot(aes(density)) +
    facet_wrap(~ central) +
    geom_histogram() + 
    ggtitle("Non Central (0) vs Central Density (1)")
```

Urban areas tend to have higher densities, as expected. Thus, the density and urban variables are collinear. The other 2 variables don't show as strong of a relationship with density, and from the scatterplot matrix are not particularly correlated with crime rate. Thus, these variables will likely not be included in our first model specification. 

Taking a closer look at crime rate against population density, this does look like a promising variable to include in the first specification.
```{r}
ggplot(crime, aes(x=density, y =crmrte)) + 
  geom_point() + 
  ggtitle("Crime Rate against Population Density") + 
  xlab("Density") + 
  ylab("Crime Rate")
```

### Examining Income-related variables

Next, we notice from the scatterplot matrix that each wage variable seems to be highly correlated with each other, and there is some positive correlation with the crime rate. We investigate these closer to see any opportunities for transformation.

```{r}
crime_wage <- crime %>% 
  select(crmrte, wcon, wtuc, wtrd, wfir, wmfg, wfed, wser, wsta, wloc) %>% 
  gather(sector, wkly_wage, -crmrte)
ggplot(crime_wage, aes(x=wkly_wage, y=crmrte)) + 
  facet_wrap(~sector) + 
  geom_point() + 
  ggtitle("Crime rate against wages across each sector") + 
  xlab("Weekly Wage") + 
  ylab("Crime Rate")
#+ theme(strip.text.x = element_text(margin = margin(2,0,2,0, "cm")))
```

Here, we see that the outlier in wser affects the x-axis is distorting the x axes for the other facets. We run the same graph without wser for a clearer visual of the other wage variables.

```{r}
crime_wage <- crime %>% 
  select(crmrte, wcon, wtuc, wtrd, wfir, wmfg, wfed, wsta, wloc) %>% 
  gather(sector, wkly_wage, -crmrte)
ggplot(crime_wage, aes(x=wkly_wage, y=crmrte)) + 
  facet_wrap(~sector) + 
  geom_point() + 
  ggtitle("Crime rate against wages across each sector without wser") + 
  xlab("Weekly Wage") + 
  ylab("Crime Rate")
#+ theme(strip.text.x = element_text(margin = margin(2,0,2,0, "cm")))
```

Many of the wage variables have a slight positive relationship with crime rate. The variables wtrd, wmfg, wfed, and (to a lesser degree) wsta and wloc seem to have a good amount of correlation with crime rate. As shown below, wser seems to have a very slight correlation with crime rate (log used because of the outlier)

```{r}
ggplot(crime, aes(x=log(wser), y=crmrte)) + geom_point()
```


These relationships explain the phenomenon that more burglaries / thefts / kidnappings are likely to be targeted on wealthier victims, so areas with higher incomes would end up having higher crime rates. 


The one point that stands out is the outlier in wser (service industry wage) when plotted against the crime rate. This reflects a county that has a substantially high wage for service workers, and has a lower crime rate. This is likely an area that has been highly gentrified and is predominantly populated my members of in service roles (with fewer individuals who are in the other, lower paying industries.) We don't believe that removing this outlier is valid, as this county could have another attribute that is worth investigating as a case study of a "successful" community with a low crime rate, and would be of high interest to our political campaign.

Taking a closer look at this outlier:

```{r}
crime_outlier <- crime %>% filter(wser>2000)
head(crime_outlier)
```

Interestingly, the tax revenue per capita is lower than what we would expect for such a supposedly affluent county. Further investigation is required here to better understand the exact job market of this population. It is likely that there are only 1 or 2 members of this county who have high paying jobs, driving up wser. In this dataset, we would hope to see a percentage breakdown of workers in each sector. This would allow us to weight each wage parameter accordingly and provide context as to how much of an influence we would expect a sector's wage to have on its crime rate.

Across all the plots, we see that points are clustered along the lower end of the x axis. Therefore, we take the log of the wage parameters. 

```{r}
ggplot(crime_wage, aes(x=log(wkly_wage), y=crmrte)) + 
  facet_wrap(~sector) + 
  geom_point() + 
  ggtitle("Crime rate against log(wages) across each sector") + 
  xlab("Log(Weekly Wage)") + 
  ylab("Crime Rate")
```
This distributes the data much more effectively. Note that even though the wage parameters may not make it into the first specification, in the later specifications we will eventually be factoring these variables in. At that time, we will be taking the log of the wages as shown above.

As mentioned above, we note that taxpc is related to the wages of workers in each county, as higher taxes are applied to individuals with a higher income. We notice this in the correlation matrix at the top of this report, where taxpc shows at least a light blue relationship with each of the wage variables individually (this is expected as we anticipate that taxpc is more tightly correlated with a linear combination of these variables). From the correlation matrix, we expect to see the same positive relationship between taxpc and crmrte. Taking a closer look at this relationship, we can verify that this relationship holds.
```{r}
ggplot(crime, aes(x=taxpc, y =crmrte)) + 
  geom_point() + 
  ggtitle("Crime Rate against Tax Per Capita") + 
  xlab("Tax Revenue per Capita") + 
  ylab("Crime Rate")
```

### Investigating demographic variables

We notice a variable pctmin80, which is the percentage of minority groups in the population. We predict that neighborhoods with higher percent minorities had lower tax revenue per capita, as socio-economic barriers often forced minority groups to take lower paying roles, and racism factors often implied that minority groups would be paid less for the same jobs as white coworkers. 

```{r}
ggplot(crime, aes(x=pctmin80, y =taxpc)) + 
  geom_point() + 
  ggtitle("Percent Minority against Tax Per Capita") + 
  xlab("Percent Minority") + 
  ylab("Tax Per Capita")
```
Contrary to the above discussion, Tax Revenue per Capita does not seem to be related to the percentage of minorities in a population. 

Now looking at percent young male. This variable is of interest as the perpetrators of crime are often thought to come from this demographic group.

```{r}
ggplot(crime, aes(x = pctymle, y = crmrte)) + 
  geom_point() + 
  ggtitle("Crime Rate against Percentage Young Males") + 
  xlab("Percent Young Male") + 
  ylab("Crime Rate")
```

### The influence of fear - probability of punishment

Now looking at the probabilities associated with arrest, conviction and prison sentence. These 3 probabilites all illustrate the likelihood of being punished for a crime. Therefore, we expect that only one of these parameters is necessary to include in our model to avoid any confounding effects.

```{r}
crime_prob_punishment <- crime %>% 
  select(crmrte, prbarr, prbconv, prbpris) %>% 
  gather(punishment, probability, -crmrte)
ggplot(crime_prob_punishment, aes(x=probability, y=crmrte)) + 
  facet_wrap(~punishment) + 
  geom_point() + 
  ggtitle("Crime rate against Probability of Arrest, Conviction, and Prison Sentence") + 
  xlab("Probability") + 
  ylab("Crime Rate")
```

As suggested by the correlation matrix and the analysis above, there is a negative relationship between the probability of arrest and conviction with crime rate.

From intuition, police presence can either be positively related with crime (more police are needed in more crime active areas) or they can be negatively related (higher plice presence serves as a deterrant of crime). In the former case, police presence is an outcome variable of crime, and in the latter case, crime is the outcome variable.

```{r}
ggplot(crime, aes(x = polpc, y = prbarr)) + 
  geom_point() + 
  ggtitle("Police Presence's effect on the Probability of Arrest") + 
  xlab("Police per Capita") + 
  ylab("Probability of Arrest")
```
There is one outlier where a very high police per capita leads to a high probability of arrest. This likely could be a neighborhood where crime is espcially high, so police are typically stationed here. This is not grounds to remove the outlier from our analysis, though we do note that, aside from this point, there doesn't seem to be a relationship between the police per capita and probability of arrest. 

We question if more police officers could be a deterrent of crime from occuring in the first place. Thus, let us plot police per capita against the crime rate. 

```{r}
ggplot(crime, aes(x = polpc, y = crmrte)) + 
  geom_point() + 
  ggtitle("Crime Rate against Police Per Capita") + 
  xlab("Police per Capita") + 
  ylab("Crime Rate")
```
From above, the more police there are the more crime there is, with the exception of the outlier at the bottom right, reflecting an area with high police per capita and low crime. This is the same point as the outlier of the previous graph. Therefore, this one county has a High number of Police with a high likelihood of arreset, and a low crime rate. This likely is an area that is actively cracking down on crime. Overall though, we cannot make any conclusions on whether police presence reduces crime, or whether crime increases police presence. Therefore, this variable will likely not be included as part of our first model specification.

```{r}
crime_outlier2 <- crime %>% filter(polpc>0.0075)
head(crime_outlier2)
```


### Bringing our analysis together

```{r}
model1 <- lm(crmrte ~ taxpc + density + pctymle + prbarr, data = crime)
model2 <- lm(crmrte ~ taxpc + density + pctymle + prbarr + prbconv + polpc + pctmin80, data = crime)
model3 <- lm(crmrte ~ taxpc + density + pctymle + prbarr + prbconv + 
               prbpris + avgsen + polpc + pctmin80 + log(wcon) + 
               log(wtuc) + log(wtrd) + log(wfir) + log(wser) + log(wmfg) + 
               log(wfed) + log(wsta) + log(wloc) + mix, data = crime)
```

# The Regression Table

crm rate as a function of Density, all the wages, probability of arrest, pct young male

```{r, results="asis"}
stargazer(model1,model2,model3
          , type ="latex"
          , column.labels   = c("Specification 1", "Specification 2 ", "Specification 3", "Specification 4")
          , report = "vc", title = "Model Summaries"
          , keep.stat = c("rsq","adj.rsq")
          , omit.table.layout = "n",
          add.lines=list(c("AIC", round(AIC(model1),1), round(AIC(model2),1), round(AIC(model3),1)),
          c("BIC", round(BIC(model1),1), round(BIC(model2),1), round(BIC(model3),1))))
```

### Comparing Specifications

Based on the results of Akaike Information Criterion (AIC) and the Bayesian Information Criterion (BIC) on each model.It is clear from the results that the first model has the best balance between parsimony and explaining the variation in the outcome variable. However it is interesting to note that both the AIC and BIC point are lower for the third specification that contains nearly all the variables than for the second specification that seems to be more parsimonious and hence could have been expected to be a better model.

# The Omitted Variables Discussion

There are several omittied variables that would be valuable in conducting this analysis:

1. Severity of crime. Crimes can vary from being petty (jaywalking or parking in a no parking zone) to severe crimes that do warrant arrest, conviction and prison sentences (kidnapping, thefts, sexual violence). Having a parameter that indicates the severity of the crime would help differentiate the varying levels of crime and focus analysis on reducing the likelihood of harsher crimes. The crime severity would be positively correlated with the crime rate and the probability of conviction but negatively correlated with the probability of arrest and the average sentence. This may lead to a negative coefficient because of the higher magnititude of the coefficient for the probability of arrest.

2. Income gap. There are several variables that point to the affluence of a region, but we are interested in seeing the percentage of upper/middle class individuals compared to percentage of lower class. We predict that the difference in these percents would be a better indicator of crime rate. Currently, we only have the wage within each sector (it is unclear whether this wage is a median or a mean or some other aggregated measure). There also could be omitted sectors, and we don't know the relative proportion of individuals in each sector. The size of the income gap may be positively correlated with the crime rate as well as the tax revenue and wage variables for high paying sectors like service while being negatively correlated with the wage variables for low paying sectors like manufacturing. As such the size of the income gap is likely to be have a positive coefficient.

### Added three omitted variables, read through
3.Police bias. Bias among police officers in certain areas may contribute to the crime rate because of spurious arrests and convictions. This may be difficult to measure directly. We would expect police bias to be positively correlated with the probability of arrests and to a lesser extent the probability of convictions. It would also be positively correlated with the crime rate leading to the coefficient being positive.

4. Crime rate in neighbouring counties. Proximity to other areas where crime is high may have an influence on the crime rate in a particular county due to spillovers of activity. This variable may be correlated with other variables like the probability of convictions and probability of arrests as well as the outcome variable, crime rate. We would as a result expect the coefficient of police bias to be positive.

5. Size of the economy. The size of the economy for each county may be a factor. Explanations could be made for crime rate to be higher or lower in a given county depending on other counties. It would be intersting to see how the crime rate varies with the size of the economy (measured by GDP or similar measure). This would likely be positively correlated with the density and tax per capita variables as well as the wage variables and the crime rate variable.  The sign of the coefficient for this variable would be expected to be positive


## Comments on the Modeling Process
Between Specification 1 and 2, we observed an increase in the adjusted R squared value, implying that the robustness of the model did indeed increase with the addition of the prbconv, polpc, and pctmin80 variables. Looking between the second and third specification, all of the variables in the second specification more or less retained their coefficient value, suggesting that the variables in specification 2 are indeed the ones we should be focusing on, hwowever the AIC and BIC scores for the third specification were better than that for the second specification which is an indication that a reasonable amount of the variation is explained by the wage variables that were left out of the second model specification.

There are a number of omitted variables, discussed above, that are suspected to have an impact on the crime rate, or that we suspect are highly correlated with the variables available in the dataset. 

Secondly there are confounding factors we must consider. We noticed that crime rate is going up due to police presence, which may seem counterintuitive. However, more police may lead to more reports of crime, so there are a large number of unreported crimes that are potentially being missed here. This leads us to question whether the crmrte variable is really a true representation of how safe a neighborhood is, as, in some areas most crimes can go unreported either due to a lack of trust in law enforcement, or simply because victims do not want to spend the energy to report a crime. We must also consider the possibility that there is a higher police presence in some variables because there is a high crime rate. This would actually suggest that polpc is an effect of crmrate than the other way round. If this is true, then we would actually advise removing the polpc variable from our model specifications.

Also, there are a variety of economic variables that could be highly correlated with another economic parameter that is causal to crmrate. Specifically, we would suggest exploring income gaps and unemployment data in counties, as these have a clearer causal mechanism to crime rate.

## Implications for the Political Campaign

In terms of actionable steps for the political campaign, we recommend looking at the outputs of the regression model with a more critical eye. From first glance, it looks like reducing the number of police, lowering taxes, and forcing young males out of counties would be the solution. These are neither advised nor are they ethical in some cases. Instead, we recommend looking at why these explanatory variables are related to crime rate. 

Some actionable steps we can recommend are creating programs to keep young men employed and off the streets, programs to improve the relationship between civilians and police forces, and making the penalties for crimes known as to deter crime from happening in the first place. Results from a future analysis that looks at income gaps explicitly would also inform which groups of individuals require a wage increase or better employement opportunities (if any).
